---
title: "TSA Assignment 8"
author: "Christian Hilscher - 1570550"
date: "4/5/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
rm(list=ls())
library(ggplot2)
library(tidyr)

#Defining color palette
cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
############################################
# Defining all functions I need later
############################################

# Function for shifting the variables back - i.e. get e_{t-1} from e_t
shift <- function(x, lag) {
  n <- length(x)
  xnew <- rep(NA, n)
  if (lag < 0) {
    xnew[1:(n-abs(lag))] <- x[(abs(lag)+1):n]
  } else if (lag > 0) {
    xnew[(lag+1):n] <- x[1:(n-lag)]
  } else {
    xnew <- x
  }
  return(xnew)
}

spectral_dens <- function(q, lambda, Mod=0){
  i = complex(real = 0, imaginary = 1)
  
  total_l <- 0
  for (l in seq((-q), q)){
    value <- exp(i*lambda*(-l))
    total_l <- total_l + value
  }
  
  total_j <- 0
  for (j in seq((-q), q)){
    value <- exp(i*lambda*j)
    total_j <- total_j + value
  }
  
  T_c <- 1/(2*q + 1)**2 * (total_j*total_l)
  spectral_d <- T_c * 1/(2*pi)
  
  if (Mod==0){
    return(spectral_d)
  } else if (Mod==1){
    return(Mod(spectral_d))
  }
  else{
    stop('Please provide either 0 or 1 as argument for Mod')
  }
}


f_lambda <- function(lambdas, q_list){
  df <- data.frame(lambdas)
  for (q in q_list){
    
    sp_densitites <- mapply(spectral_dens, q=q, lambda=lambdas)
    num_sp_densitites <- Mod(sp_densitites)
    
    name <- "density_q"
    fullname <- paste(name, toString(q), sep="")
    
    df[[fullname]] <- num_sp_densitites
  }
  return(df)
}

VR <- function(lambdas, q_list){
  df <- data.frame(lambdas)
  for (q in q_list){
    s_lambda0 <- vector()
    for (lambda in lambdas){
      s_current <- 2* integrate(spectral_dens, 0, lambda, q=q, Mod=1)$value
      s_lambda0 <- rbind(s_lambda0, s_current)
    }
    
    s_pi <- 2* integrate(spectral_dens, 0, pi, q=q, Mod=1)$value
    
    name <- "VR_q"
    fullname <- paste(name, toString(q), sep="")
    
    results <- s_lambda0/s_pi
    df[[fullname]] <- results
  }
  
  return(df)
}

prepare_plot <- function(df){
  df_plot <- pivot_longer(df, -lambdas, names_to="q", values_to="values")
  return(df_plot)
}
```

# Question 13

I first rewrite $c_j$ as a function of the transfer function:

\begin{align*}
c_j &= \frac{1}{2 \pi} \int_{-\pi}^{\pi} e^{ij \lambda} \Psi(\lambda) d\lambda \\
&= \frac{1}{2 \pi} \left ( \underbrace{\int_{-\pi}^{- \lambda_h} (\cdot) d\lambda}_{0} + \int_{- \lambda_h}^{- \lambda_l} (\cdot) d\lambda + \underbrace{\int_{-\lambda_l}^{- \lambda_h} (\cdot) d\lambda}_{0} + \int_{\lambda_l}^{\lambda_h} (\cdot) d\lambda + \underbrace{\int_{\lambda_h}^{\pi} (\cdot) d\lambda}_{0} \right ) 
\end{align*}

This splitting up now allows me to analyse the terms which are not zero:

\begin{align*}
c_j =& \frac{1}{2 \pi} \left( \int_{- \lambda_h}^{- \lambda_l} e^{ij \lambda} d\lambda + \int_{\lambda_l}^{\lambda_h} e^{ij \lambda} d\lambda \right) \\
=& \frac{1}{2 \pi} \left( \left[ \frac{1}{ij} e^{ij \lambda} \right]^{- \lambda_l}_{- \lambda_h} + \left[ \frac{1}{ij} e^{ij \lambda} \right]^{\lambda_h}_{\lambda_l} \right) \qquad \text{for } j \neq 0 \\
=& \frac{1}{2 \pi} \frac{1}{ij} \left( e^{ij \lambda_h} - e^{- ij \lambda_h} \right) - \left( e^{ij \lambda_l} - e^{- ij \lambda_l} \right) \\
=& \frac{1}{2 \pi} \frac{1}{ij} \left( 2i \sin(j \lambda_h) \right) - \left( 2i \sin(j \lambda_l) \right) \\
c_j =& \frac{1}{\pi j} (\sin(j \lambda_h) - \sin(j \lambda_l)) \quad \text{for } j \neq 0
\end{align*}

Coming to $j=0$ one can show that

\begin{align*}
c_0 =& \frac{1}{2 \pi} \left( \int_{- \lambda_h}^{- \lambda_l} e^{0i \lambda} d\lambda + \int_{\lambda_l}^{\lambda_h} e^{0i \lambda} d\lambda \right) \\
=& \frac{1}{2 \pi} \left( \int_{- \lambda_h}^{- \lambda_l} 1 d\lambda + \int_{\lambda_l}^{\lambda_h} 1 d\lambda \right) \\
=& \frac{1}{2 \pi} \left( - \lambda_h + \lambda_h + \lambda_h - \lambda_l \right) \\
c_0 =& \frac{1}{\pi} (\lambda_h - \lambda_l)
\end{align*}

# Question 14

### DGP
To generate the data I use 500 points to represent all $\lambda_0$ which is enough to make the grpahs look nice and convey the overall idea without being computationally too intensive.

```{r}
### Initiating starting values and dataframes for storage
points <- 500
lambdas <- seq(from=0, 
               to=pi, 
               length.out = points)
q_list = c(1, 3, 10)

df_f <- f_lambda(lambdas, q_list)
df_VR <- VR(lambdas, q_list)
```

### Getting the spectral density function

As a start I rewrite the process a little bit

\begin{align*}
x_t =& \frac{1}{2q+1} \sum_{|j|\le q} \epsilon_{t-j} \\
B(L) =& \frac{1}{2q+1} \sum_{|j|\le q} \\
B(e^{-i\lambda}) =& \frac{1}{2q+1} \sum_{|j|\le q} e^{-i\lambda}
\end{align*}

Now using $\sigma = 1$, the spectral density $f_x(\lambda)$ is given by

\begin{align*}
f_x(\lambda) =& |B(e^{-i\lambda})|^2 \frac{\sigma}{2 \pi} \\
=& \frac{1}{2 \pi} \left( \frac{1}{2q+1} \sum_{|j|\le q} e^{-i\lambda} \right)\left( \frac{1}{2q+1} \sum_{|j|\le q} e^{i\lambda} \right) \\
=& \frac{1}{2 \pi} \frac{1}{(2q+1)^2} \sum_{|j|\le q} e^{-i\lambda} \sum_{|j|\le q} e^{i\lambda} \tag{1}
\end{align*}

The implementation of $(1)$ is given by the function below with the argument _Mod_ set to 0 to get imaginary numbers.

```{r}
spectral_dens <- function(q, lambda, Mod=0){
  i = complex(real = 0, imaginary = 1)
  
  # First sum
  total_l <- 0
  for (l in seq((-q), q)){
    value <- exp(i*lambda*(-l))
    total_l <- total_l + value
  }
  
  # Second sum
  total_j <- 0
  for (j in seq((-q), q)){
    value <- exp(i*lambda*j)
    total_j <- total_j + value
  }
  
  T_c <- 1/(2*q + 1)**2 * (total_j*total_l)
  spectral_d <- T_c * 1/(2*pi)
  
  # Whether or not to take the norm of the vector to convert it back
  if (Mod==0){
    return(spectral_d)
  } else if (Mod==1){
    return(Mod(spectral_d))
  }
  else{
    stop('Please provide either 0 or 1 as argument for Mod')
  }
}
```

To actually compute the spectral density for $\lambda_0 \in [0, \pi]$ I run the following function which takes the list of 500 lambdas and the 3 values of q as arguments. The function _mapply_ allows me to use the whole vector as input for the scalar valued function defined above. For plotting the values I take the norm and finally add them to a dataframe.

```{r}
f_lambda <- function(lambdas, q_list){
  df <- data.frame(lambdas)
  for (q in q_list){
    
    sp_densitites <- mapply(spectral_dens, q=q, lambda=lambdas)
    num_sp_densitites <- Mod(sp_densitites)
    
    name <- "density_q"
    fullname <- paste(name, toString(q), sep="")
    
    df[[fullname]] <- num_sp_densitites
  }
  return(df)
}
```

The result of this can be seen in the first figure

```{r, include=FALSE}
df_f_plot <- prepare_plot(df_f)

p1=ggplot(data=df_f_plot, aes(x=lambdas, y=values, colour=q)) + 
  geom_line(alpha=0.8, size=0.7) + 
  scale_colour_manual(values = cbPalette) +
  xlab('Lambda') + 
  ylab(expression(f[x](lambda))) + 
  ggtitle("Values of the spectal density function over the interval [0,pi]")
```

```{r, echo=FALSE}
plot(p1)
```

### Getting the variance shares

For getting the share of the varainces whose components happen with a frequency lower than $\lambda_0$, I run the following function. Similar to the spectral density function it takes as arguments the list of lambdas and the different values of _q_. Although all of the results of the integration have only real parts, I need to convert them since R still thinks of them as complex values and then the integration function does not work. At the end again a dataframe is being filled up since this way of storing the values makes plotting easier afterwards.

```{r}
VR <- function(lambdas, q_list){
  df <- data.frame(lambdas)
  for (q in q_list){
    s_lambda0 <- vector()
    for (lambda in lambdas){
      s_current <- 2* integrate(spectral_dens, 0, lambda, q=q, Mod=1)$value
      s_lambda0 <- rbind(s_lambda0, s_current)
    }
    
    s_pi <- 2* integrate(spectral_dens, 0, pi, q=q, Mod=1)$value
    results <- s_lambda0/s_pi
    
    name <- "VR_q"
    fullname <- paste(name, toString(q), sep="")
    
    df[[fullname]] <- results
  }
  
  return(df)
}
```

Talking about plotting, $VR(\lambda_0; q)$ is 

```{r, include=FALSE}
df_VR_plot <- prepare_plot(df_VR)

p2=ggplot(data=df_VR_plot, aes(x=lambdas, y=values, colour=q)) + 
  geom_line(alpha=0.8, size=0.7) + 
  scale_colour_manual(values = cbPalette) +
  xlab('Lambda') + 
  ylab(expression(VR(lambda[0],q))) +
  ggtitle("Values of the VR as a function of lambda between [0,pi]")
```


```{r, echo=FALSE}
plot(p2)
```